### 2016-03-17 text/MS.Rmd
new version of MS with corrected dataset

```{r, results="hide", warnings=FALSE, echo=FALSE, message=FALSE}
load('../output/RDatafiles/analyses.RData')
load('../output/RDatafiles/numbers.RData')
source('../R/functions-analyses.R')
source('../R/functions-figures.R')
cleanbib()
myRefs   <-  listBibs()
exporBibs(myRefs, erase=TRUE)
```
**Temperature effects on mass-scaling exponent of metabolic rates: a manipulative test**

@@

1. Organismal metabolic rate often scales allometrically with body mass, both inter and intra-specifically. Metabolic rates are also known to increase with temperature, but most theoretical tests of mass scaling and temperature dependence have treated both phenomena independently.  
2. In the present study, we use a manipulative approach to test for an association between metabolic mass scaling and temperature dependence in four species belonging to two phyla of colonial marine invertebrates.  
3. We show that mass-scaling exponents and temperature dependencies interact in different ways for different species. In so doing, we discuss the potentials of our results to multi-level scaling theory, showing that some species might be winners while others might be losers under different temperature scenarios.  
4. These findings highlight the need for new metabolic theory that formally incorporate interactions between mass-scaling and temperature dependence in ectothermic organisms, in order to better explain documented variation in the mass-scaling exponent of metabolic rate, and lead to novel and testable predictions for associations between individual metabolism and community structure.

@@

Metabolic theories provide a potential to unite hierarchical levels of biological organisation, from individuals to ecosystems `r myCite(c('nisbet2000janimecol', 'brown2004', 'vandermeer2006tree'))`. Although definitive tests of their underlying assumptions remain rare `r myCite(c('kearney2012amnat', 'newberry2015ploscomputbiol'))`, myriad studies examined their predictions with regards to the relationship between individual level metabolic rates, $B_i$, and individual body mass, $M_i$ `r myCite(c('kleiber1932', 'kleiber1961', 'peters1983', 'schmidtnielsen1984', 'glazier2005biolrev', 'delong2010', 'glazier2010biolrev'))`. The relationship between $B_i$ and $M_i$ is typically described by a power function of the form $B_i = B_oM_i^{\alpha}$, where $B_o$ is a normalisation constant independent of mass and $\alpha$ is a dimensionless scaling exponent. Numerous studies have demonstrated that, for most animals, individual metabolic rate scales sub-linearly with body mass, with $\alpha$ estimates varying with many biotic and abiotic variables (e.g. `r myCite(c('lovegrove2000amnat', 'white2009evolution', 'killen2010ele', 'kolokotrones2010nature', 'vaca2010biollett', 'glazier2011ecolmongr', 'ketola2012jlinsoc', 'white2012procb', 'white2013jcompphysb', 'barneche2014ele'))`), generally falling between 0.5 and 1, and averaging around 0.75. While these studies demonstrate clearly that the value of $\alpha$ is not fixed, they each share a common limitation: each study tests for an association between $\alpha$ and a biotic or abiotic variable, using a value of $\alpha$ that is estimated from the association between $B_i$ and $M_i$ for individuals or species of different size. Such correlative studies cannot address the causal association between $B_i$ and $M_i$, because $M_i$ is correlated with a range of other environmental, physiological, and life history variables (see e.g. `r myCite(c('peters1983', 'calder1984book', 'schmidtnielsen1984'))` for comprehensive reviews) and many of these variables are also associated with $B_i$ `r myCite(c('mcnab2002book', 'konarzewski2013jcompphysb', 'white2013jcompphysb'))`.

In contrast to studies that make use of intact individual organisms, colonial animals provide an opportunity for more definitive tests of metabolic theories. Colony size can be manipulated experimentally, hence allowing for the potential to remove the confounding effects of body size and life-history traits (e.g. `r myCite('white2011amnat')`). Furthermore, different colonial species may grow in different dimensions, and the mass-scaling exponent $\alpha$ is explicitly predicted to change with number of body dimensions `r myCite(c('nakaya2005procb', 'white2011amnat', 'kearney2012amnat'))`. For example, the fractal geometry model `r myCite(c('west1997science', 'west1999science'))` predicts that $\alpha = n/(n + 1)$ for a $n$-dimensional organism, hence a value of $\alpha$ = 0.75 is predicted for three-dimensional animals while a value of $\alpha$ = 0.67 is predicted for two-dimensional animals (but see `r myCite(c('enquist2009nature', 'savage2008ploscompbiol', 'banavar2010pnas', 'kolokotrones2010nature', 'price2007pnas', 'koontz2009amjbot'))`). Moreover, experimental size manipulation in colonial organisms allows for tests of how ecological processes affect metabolic rates. For instance, it has been demonstrated that mimicking partial predation through size manipulation yields different responses in growth and reproductive onset in different species of encrusting and arborescent bryozoans `r myCite(c('bone2005jembe', 'hart2009ecologyb'))`. However, to our knowledge, few studies have explicitly investigated the effects of partial predation on individual metabolic rates (e.g. `r myCite('white2011amnat')`).

Metabolic rates are also influenced by temperature, typically following an exponential function `r myCite(c('gillooly2001', 'kooijman2009book'))`. Existing metabolic theories assume nevertheless different paths through which temperature affects metabolic rates. For example, the fractal geometry model `r myCite(c('west1997science', 'west1999science'))` assumes that temperature affects the metabolic normalisation only `r myCite(c('gillooly2001', 'brown2004'))`, i.e. mass-scaling and temperature dependence of metabolic rates are fundamentally independent; contrarily, the Dynamic Energy Budget (DEB) theory for metabolic organisation `r myCite('kooijman2009book')` predicts changing relationships between mass scaling and temperature dependence through the effects of temperature on assimilation, maintenance, growth, and maturation `r myCite('kooijman2009book')`. Whether temperature dependence and mass scaling truly interact remains a contentious issue, though recent empirical evidence indicates that they do, at least in fishes `r myCite('killen2010ele')`. Testing such interaction is of great ecological significance because different mass-scaling exponents under different temperatures would potentially indicate contrasting life-history trajectories `r myCite('sebens1987annurevecolsys')` for a given species under different environments **(REFS??)**.

@@

*1. Animal collections*  
To collect our study organisms, roughened acetate sheets secured to the underside of 6 mm thick PVC sheets suspended from floating pontoons at a depth of 1 m were deployed at Manly Boat Harbour, Queensland, Australia (27˚27' S, 153˚11' E) and checked regularly to remove fouling organisms other than our species of interest: *Bugula neritina* and *B. stolonifera* (arborescent bryozoan), *Hippopodina* sp. (encrusting bryozoan), and an unidentified species of sponge belonging to the Microcionidae family `r myCite('hart2009ecology')` (Table 1). After 6 weeks in the field, sheets bearing colonies were returned to the University of Queensland where they were maintained in aerated seawater for up to 48 h prior to measurements. For encrusting colonies, sections of acetate bearing whole colonies were then cut from the sheets and fragmented (see `r myCite('white2011amnat')` for details and tests of potential experimental artefacts). Arborescent colonies were cut directly from the acetate sheets and fragmented.

*2. Respiration measurements*  
Individual metabolic rates (**provide units**, *n* = `r nrow(metRates)`), $B_i$, were measured as rate of oxygen consumption, $\dot{V}\textrm{O}_\textrm{2}$ of colony fragments was measured at 10 or 25 ˚C using a 144-channel PreSens Sensor Dish Reader (SDR, AS-1 Scientific, Wellington, New Zealand) according to standard techniques `r myCite(c('koster2008meps', 'lighton2008book', 'white2011amnat'))`. $\dot{V}\textrm{O}_\textrm{2}$ was measured by placing the colony or fragment in a glass vial containing 0.2$\mu$m filtered sea water and a non-consumptive $\textrm{O}_\textrm{2}$ sensor spot and calculated from the rate of change of $\textrm{O}_\textrm{2}$ saturation ($m_a$, \% h^‑1^) as:

@@

where $m_b$ is the rate of change of $\textrm{O}_\textrm{2}$ saturation for blank vials containing no bryozoans (\% h^-1^), ${\beta}_{\textrm{O}_\textrm{2}}$ is the oxygen capacitance of air-saturated sea water at 10 ˚C (6.26 ml l^-1^) or 24 ˚C (4.615 ml l^-1^) [**double-check these numbers. Shouldn't the second be 25 ˚C?**] `r myCite('cameron1986book')` and $V$ is water volume (chambers were 0.005 l in volume and water volume was calculated by subtracting the volume of acetate and animals). At least four blank vials were recorded simultaneously with each 24-well plate to account for microbial oxygen consumption and sensor spots were calibrated with air-saturated seawater (100\% A.S.) and water containing 2\% sodium sulfite (0\% A.S.). Following measurement of $\dot{V}\textrm{O}_\textrm{2}$, colonies or fragments were blotted dry and weighed to 0.1 mg (Sartorius A 200 S, Sartorius AG, Göttingen, Germany). All measurements of $\dot{V}\textrm{O}_\textrm{2}$ were made in one of four dark constant temperature cabinets, with temperatures within a run randomised among cabinets.

@@

Both models were fitted in a Bayesian framework by calling *JAGS* version 3.4.0 from the R package *R2jags* version 0.05-03 `r myCite('r2jags')` in order to derive posterior distributions and associated 95\% credible intervals (CIs) for the fitted parameters. Random effects were assumed to be normally distributed, with means of 0. Fitted parameters were assigned priors that were vague (i.e. locally uniform over the region supported by the likelihood) `r myCite('kruschke2014book')`. The posterior distributions of model parameters were estimated using Markov chain Monte Carlo (MCMC) methods by constructing five chains of 1,000,000 steps each, including 500,000-step burn-in periods. Chains were thinned using a 250-step interval, so a total of 10,000 steps were retained to estimate posterior distributions (i.e. 5 $\times$ (1,000,000 - 500,000)/250 = 10,000). Each nested model was compared against the full model using leave-one-out cross-validation (LOO), which, similar to widely applicable information criterion (WAIC), is a fully Bayesian model selection procedure for estimating pointwise out-of-sample prediction accuracy using the log-likelihood evaluated at the posterior simulations of the parameter values `r myCite('hooten2015ecology')`. Model significance was assessed by calculating a *z*-score from the average difference $\pm$ SE between expected log pointwise predictive density of both models, and models were considered significantly different at $\alpha \le$ 0.05. Model comparison was implemented using the R package *loo* version 0.1.3 `r myCite('loo')`. All data and code necessary to reproduce this paper, its analyses, tables and figures can be obtained on GitHub https://github.com/dbarneche/MTRBrEs.

@@

Our model selection approach indicated that the three-way interaction model (equation 2) performed significantly better than all nested two- and one-way interaction models (Table 2), thus implying that different species present different interactions between mass-scaling and temperature dependence. Particularly, the best model yields no interaction between mass scaling and temperature dependence for *Bugula stolonifera* and the Microcionidae sponge (large overlap between 95\% credible intervals for the mass-scaling slopes, $\alpha$) (Figure 1). Still, while the mass-scaling slopes are statistically the same between temperatures for these species, there is evidence for higher metabolic normalisations (i.e. intercepts), $B_o$, at higher temperatures (though not statistically significant for Microcionidae). For example, metabolic rates increase `r rounded(exp(averageEstimates$lnB25[averageEstimates$species == 'Stolonifera'])/exp(averageEstimates$lnB10[averageEstimates$species == 'Stolonifera']))`-fold between 10 ˚C and 25 ˚C for *B. stolonifera*, yielding a $Q_{10}$ of `r rounded(q10andErs$q10[q10andErs$species == 'Stolonifera'])`. It is worth noting that this $Q_{10}$ value is equivalent to an activation energy of `r rounded(q10andErs$er[q10andErs$species == 'Stolonifera'], 2)` eV (i.e. $E_r = \frac{k \times 283.15 \times 298.15 \times 1.5 \times \textrm{ln}Q_{10}}{(298.15 - 283.15)}$, where $E_r$ is an activation energy and $k$ is the Boltzmann constant, 8.62 $\times$ 10^-5^ eV K^-1^), which is consistent with the average 06 -- 0.7 eV reported in the Metabolic Theory of Ecology literature `r myCite(c('gillooly2001','yvondurocher2012a'))`. Different from the above species, the best model indicates that *Bugula neritina* and *Hippopodina* sp. present opposing interactions between mass-scaling and temperature dependence (Figure 1). For example, average posterior parameter estimates indicate that the difference in metabolic rates between 25 ˚C and 10 ˚C for *Hippopodina* sp. decreases, on average, `r rounded(tempDeltas[1,'tempDelta'])`-fold moving from `r round(exp(2))` to `r round(exp(5))` mg. For *Bugula neritina*, however, rates increase `r rounded(1/tempDeltas[3,'tempDelta'])`-fold along the same body-mass gradient.

Posterior distributions for mass-scaling exponents show varied consistency with predictions from the two metabolic theories. For example, *Hippopodina* sp. shows exponents that encompass both 0.67 (Fractal geometry model prediction for 2-D organisms; `r myCite('koontz2009amjbot')`) and 0.75 (Fractal geometry model prediction for 3-D organisms; `r myCite(c('west1997science', 'west1999science'))`) at 10 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's10_97.5'], 2)`), but that are significantly shallower than 0.75 at 25 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's25_97.5'], 2)`). On the other hand, *Bugula neritina* presents scaling exponents that overlap both 0.67 and 0.75 at 25 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's25_97.5'], 2)`), but are significantly shallower than these values at 10 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's10_97.5'], 2)`). Microcionidae presents exponents that only overlap 0.67 at 10 ˚C, but not at 25 ˚C (95\% CI at 10 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's10_97.5'], 2)`; 95\% CI at 25 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's25_97.5'], 2)`). Finally, *Bugula stolonifera* presents values that overlap both 0.67 and 0.75 at both temperatures (95\% CI at 10 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's10_97.5'], 2)`; 95\% CI at 25 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's25_97.5'], 2)`).

@@

Here, we show experimentally that the mass-scaling of metabolic rates changes between two extreme temperatures, and in different ways across different species of marine invertebrates. We build on recent empirical evidence documenting similar changes (e.g. `r myCite('killen2010ele')`), yielding further support to the need for new metabolic theories (or reformulation of current ones) that formally incorporate interactive effects between mass scaling and temperature dependence of biological rates. Our results, however, do not indicate consistent mass-temperature interactions across growth forms. Possibly, another important axis involved, which has not been considered here, is the acclimation capacity of species upon acute temperature changes (e.g. `r myCite('rezende2014functecol')`). Our findings are constrained to acute-related changes, and therefore do not consider temporal acclimation or even adaptation to changes in environmental temperature.

The fact that $B_o$ and $\alpha$ covary in different ways depending on species and temperature (Figure 1) suggests that the metabolic costs of maintenance will vary for different species at different environmental temperatures. For example, *Bugula neritina* presented a shallower mass-scaling exponent at colder temperatures, possibly indicating that large cold-environment individuals are more efficient in spending energy than counterparts in warm environments. Thus, for this species, all else being equal, increases in temperature decrease the metabolic benefits of attaining a larger size while decreases in temperature reduce the metabolic normalisation. On the other hand, our results indicate that *Hippopodina* sp. presents a reduction in metabolic benefits of attaining a larger size at colder temperatures. Thus, for this encrusting species, one could predict that individuals living at higher temperatures should allocate more resources to attaining larger sizes (as opposed to reproduction, for instance) while individuals living at lower temperatures will be under less pressure to allocate resources to size. Importantly, though, we note that there are many extrinsic factors that may affect abundance and selection on body size (e.g. predation, competition, etc; `r myCite('barneche2016procb')`). Yet, such differentials in the costs and benefits of attaining large size may help explain why some species living in cold temperatures present a slower pace of life (e.g. `r myCite('rosa2010jmba')`). Moreover, they may represent necessary improvements on current attempts to scale up energy from individuals to higher levels of organisation (e.g. `r myCite(c('allen2005','yvondurocher2012b','barneche2014ele','barneche2016procb'))`).

>>>>

```{r, results="hide", warnings=FALSE, echo=FALSE, message=FALSE}
load('../output/RDatafiles/analyses.RData')
load('../output/RDatafiles/numbers.RData')
source('../R/functions-analyses.R')
source('../R/functions-figures.R')
cleanbib()
myRefs   <-  listBibs()
exporBibs(myRefs, erase=TRUE)
```
**Temperature effects on mass-scaling exponents in colonial animals: a manipulative test**

@@

1. Metabolic theory posits that size and temperature are fundamental drivers of ecological patterns and processes, via their respective associations with metabolic rates. Metabolic rates are well known to increase allometrically with body mass and approximately exponentially with temperature, but most applications of metabolic theory treat these effects as independent.  
2. Almost all studies of intraspecific scaling of unitary organisms must rely on ontogenetic differences in size to examine its relationship with metabolic rate, thereby potentially confounding age, nutrition or growth rate with size –- all of which can affect metabolic rate. In the present study, we use a manipulative approach to test for an association between metabolic mass scaling and temperature dependence in four species of colonial marine invertebrates belonging to two phyla.  
3. While increased temperature increased metabolic rate, we found no consistent effect of temperature or growth form on metabolic scaling; rather the effects were species-specific. We show that mass-scaling exponents and temperature dependencies interact in ways that are unanticipated by extant theory.  
4. The allometric scaling of metabolic rate that we observe in the present study contrasts with the isometric scaling of water clearance and feeding rates observed in some colonial marine invertebrates. Our findings suggest that the often observed competitive superiority of colonial organisms over unitary ones may arise because energy availability for growth and or reproduction (the difference between energy intake and expenditure) increases more strongly with size in colonial organisms.

@@

Metabolic theories seek to unite hierarchical levels of biological organisation, from individuals to ecosystems `r myCite(c('nisbet2000janimecol', 'brown2004', 'vandermeer2006tree'))`. Although definitive tests of their underlying assumptions remain rare `r myCite(c('kearney2012amnat', 'newberry2015ploscomputbiol'))`, myriad studies examined their predictions with regards to the relationship between individual level metabolic rates, $B_i$, and individual body mass, $M_i$ `r myCite(c('kleiber1932', 'kleiber1961', 'peters1983', 'schmidtnielsen1984', 'glazier2005biolrev', 'delong2010', 'glazier2010biolrev'))`. The relationship between $B_i$ and $M_i$ is typically described by a power function of the form $B_i = B_oM_i^{\alpha}$, where $B_o$ is a normalisation constant independent of mass and $\alpha$ is a dimensionless scaling exponent. Numerous studies have demonstrated that, for most animals, individual metabolic rate scales sub-linearly with body mass, with $\alpha$ estimates varying with many biotic and abiotic variables (e.g. `r myCite(c('lovegrove2000amnat', 'white2009evolution', 'killen2010ele', 'kolokotrones2010nature', 'vaca2010biollett', 'glazier2011ecolmongr', 'ketola2012jlinsoc', 'white2012procb', 'white2013jcompphysb', 'barneche2014ele'))`), generally falling between 0.5 and 1. While these studies demonstrate clearly that the value of $\alpha$ is not fixed, they each share a common limitation: each study tests for an association between $\alpha$ and a biotic or abiotic variable, using a value of $\alpha$ that is estimated from the association between $B_i$ and $M_i$ for individuals or species of different size. Such correlative studies cannot address the causal association between $B_i$ and $M_i$, because $M_i$ is correlated with a range of other environmental, physiological, and life history variables (see e.g. `r myCite(c('peters1983', 'calder1984book', 'schmidtnielsen1984'))` for comprehensive reviews) and many of these variables are also associated with $B_i$ `r myCite(c('mcnab2002book', 'konarzewski2013jcompphysb', 'white2013jcompphysb'))`.

In contrast to studies that make use of intact individual organisms, colonial animals provide an opportunity for more definitive tests of metabolic theories. Colony size can be manipulated experimentally, hence allowing for the potential to remove the confounding effects of body size and life-history traits (e.g. `r myCite('white2011amnat')`). Furthermore, different colonial species may grow in different dimensions, and the mass-scaling exponent $\alpha$ is explicitly predicted to change with number of body dimensions `r myCite(c('nakaya2005procb', 'white2011amnat', 'kearney2012amnat'))`. For example, the fractal geometry model `r myCite(c('west1997science', 'west1999science'))` predicts that $\alpha = n/(n + 1)$ for a $n$-dimensional organism, hence a value of $\alpha$ = 0.75 is predicted for three-dimensional animals while a value of $\alpha$ = 0.67 is predicted for two-dimensional animals (`r myCite(c('enquist2009nature', 'savage2008ploscompbiol', 'banavar2010pnas', 'kolokotrones2010nature', 'koontz2009amjbot'))`; but see `r myCite('price2007pnas')` for additional predictions of $\alpha$ upon modifications of canonical assumptions). Moreover, experimental size manipulation in colonial organisms allows for tests of how ecological processes such as predation or disturbance affect metabolic rates. For instance, it has been demonstrated that mimicking the effects of partial predation of physical disturbance through experimental size manipulation yields different responses in growth and reproductive onset in different species of encrusting and arborescent bryozoans `r myCite(c('bone2005jembe', 'hart2009ecologyb'))`. Yet, the metabolic consequences of these size manipulations remains unclear.

Metabolic rates are influenced by temperature, typically following an exponential function `r myCite(c('gillooly2001', 'kooijman2009book'))`. Existing metabolic theories assume nevertheless different paths through which temperature affects metabolic rates. For example, the fractal geometry model `r myCite(c('west1997science', 'west1999science'))` assumes that temperature affects the metabolic normalisation only `r myCite(c('gillooly2001', 'brown2004'))`, i.e. mass-scaling and temperature dependence of metabolic rates are fundamentally independent. In contrast, the Dynamic Energy Budget (DEB) theory for metabolic organisation `r myCite('kooijman2009book')` predicts changing relationships between mass scaling and temperature dependence through the effects of temperature on assimilation, maintenance, growth, and maturation `r myCite('kooijman2009book')`. Whether temperature dependence and mass scaling interact remains a contentious issue, though recent empirical evidence indicates that they do, at least in fishes `r myCite('killen2010ele')`. Testing such interactions matters because different mass-scaling exponents under different temperatures yield different metabolic efficiencies associated with different body sizes. For example, a body size that maximizes scope for reproduction under one temperature may be suboptimal in a cooler temperature or vice versa `r myCite('sebens1987annurevecolsys')`.

@@

*1. Animal collections*  
To collect our study organisms, roughened acetate sheets secured to the underside of 6 mm thick PVC sheets suspended from floating pontoons at a depth of 1 m were deployed at Manly Boat Harbour, Queensland, Australia (27˚27' S, 153˚11' E) and checked regularly to remove fouling organisms other than our species of interest: *Bugula neritina* and *B. stolonifera* (arborescent bryozoans that have a three dimensional tree-like growth form), *Hippopodina* sp. (an encrusting bryozoan that grows mostly in two dimensions as flat disc), and an unidentified species of sponge belonging to the Microcionidae family (`r myCite('hart2009ecology')`; again a flat species that grows mostly in two dimensions) (Table 1). After 6 weeks in the field, sheets bearing colonies were returned to the University of Queensland where they were maintained in aerated seawater for up to 48 h prior to measurements. For encrusting colonies, sections of acetate bearing whole colonies were then cut from the sheets and fragmented (see e.g. `r myCite(c('nakaya2005procb', 'white2011amnat'))` for details and tests of potential experimental artefacts). Arborescent colonies were cut directly from the acetate sheets and fragmented.

*2. Respiration measurements*  
Individual metabolic rates (converted to $\mu$l $\textrm{O}_\textrm{2}$ h^-1^, *n* = `r nrow(metRates)`), $B_i$, were measured as rate of oxygen consumption, $\dot{V}\textrm{O}_\textrm{2}$ of colony fragments was measured at nominal temperatures of 10 or 25 ˚C using a 144-channel PreSens Sensor Dish Reader (SDR, AS-1 Scientific, Wellington, New Zealand) according to standard techniques `r myCite(c('koster2008meps', 'lighton2008book', 'white2011amnat'))`. $\dot{V}\textrm{O}_\textrm{2}$ was measured by placing the colony or fragment in a glass vial containing 0.2$\mu$m filtered sea water and a non-consumptive $\textrm{O}_\textrm{2}$ sensor spot and calculated from the rate of change of $\textrm{O}_\textrm{2}$ saturation ($m_a$, \% h^‑1^) as:

@@

where $m_b$ is the rate of change of $\textrm{O}_\textrm{2}$ saturation for blank vials containing no bryozoans (\% h^-1^), ${\beta}_{\textrm{O}_\textrm{2}}$ is the oxygen capacitance of air-saturated sea water at 10 ˚C (6.4 ml l^-1^) or 25 ˚C (4.74 ml l^-1^) `r myCite('cameron1986book')` and $V$ is water volume (chambers were 0.005 l in volume and water volume was calculated by subtracting the volume of acetate and animals). At least four blank vials were recorded simultaneously with each 24-well plate to account for microbial oxygen consumption and sensor spots were calibrated with air-saturated seawater (100\% A.S.) and water containing 2\% sodium sulfite (0\% A.S.). Following measurement of $\dot{V}\textrm{O}_\textrm{2}$, colonies or fragments were blotted dry and weighed to 0.1 mg (Sartorius A 200 S, Sartorius AG, Göttingen, Germany). All measurements of $\dot{V}\textrm{O}_\textrm{2}$ were made in one of four dark constant temperature cabinets, with temperatures within a run randomised among cabinets.

@@

All models were fitted in a Bayesian framework by calling *JAGS* version 3.4.0 from the R package *R2jags* version 0.05-03 `r myCite('r2jags')` in order to derive posterior distributions and associated 95\% credible intervals (CIs) for the fitted parameters. Random effects were assumed to be normally distributed, with means of 0. Fitted parameters were assigned priors that were vague (i.e. locally uniform over the region supported by the likelihood) `r myCite('kruschke2014book')`. The posterior distributions of model parameters were estimated using Markov chain Monte Carlo (MCMC) methods by constructing five chains of 1,000,000 steps each, including 500,000-step burn-in periods. Chains were thinned using a 250-step interval, so a total of 10,000 steps were retained to estimate posterior distributions (i.e. 5 $\times$ (1,000,000 - 500,000)/250 = 10,000). Each nested model was compared against the full model using leave-one-out cross-validation (LOO), which, similar to widely applicable information criterion (WAIC), is a fully Bayesian model selection procedure for estimating pointwise out-of-sample prediction accuracy using the log-likelihood evaluated at the posterior simulations of the parameter values `r myCite('hooten2015ecology')`. Model significance was assessed by calculating a *z*-score from the average difference $\pm$ SE between expected log pointwise predictive density of both models, and models were considered significantly different at $\alpha \le$ 0.05. Model comparison was implemented using the R package *loo* version 0.1.3 `r myCite('loo')`. All data and code necessary to reproduce this paper, its analyses, tables and figures can be obtained on GitHub https://github.com/dbarneche/MTRBrEs.

@@

Our model selection approach indicated that the three-way interaction model (equation 2) performed significantly better than all nested two- and one-way interaction models (Table 2), thus implying that different species present different interactions between mass-scaling and temperature dependence. Particularly, the best model yields no interaction between mass scaling and temperature dependence for *Bugula stolonifera* and the Microcionidae sponge (large overlap between 95\% credible intervals for the mass-scaling slopes, $\alpha$) (Figure 1). Still, while the mass-scaling slopes are statistically the same between temperatures for these species, there is evidence for higher metabolic normalisations (i.e. intercepts), $B_o$, at higher temperatures (though not statistically significant for Microcionidae). For example, metabolic rates increase `r rounded(exp(averageEstimates$lnB25[averageEstimates$species == 'Stolonifera'])/exp(averageEstimates$lnB10[averageEstimates$species == 'Stolonifera']))`-fold between 10 ˚C and 25 ˚C for *B. stolonifera*, yielding a $Q_{10}$ of `r rounded(q10andErs$q10[q10andErs$species == 'Stolonifera'])`, which is equivalent to an activation energy of `r rounded(q10andErs$er[q10andErs$species == 'Stolonifera'], 2)` eV `r myCite(c('gillooly2001','yvondurocher2012a'))`. The best model indicates that the difference in metabolic rates between 25 ˚C and 10 ˚C for *Hippopodina* sp. decreases, on average, `r rounded(tempDeltas[1,'tempDelta'])`-fold moving from `r round(exp(2))` to `r round(exp(5))` mg. For *Bugula neritina*, however, rates increase `r rounded(1/tempDeltas[3,'tempDelta'])`-fold along the same body-mass gradient.

Average posterior estimates of mass-scaling exponents are inconsistent among the studied species and between growth forms (2D and 3D). For example, for organisms that exhibit 2D growth, *Hippopodina* sp. shows exponents that encompass both 0.67 (Fractal geometry model prediction for 2D organisms; `r myCite('koontz2009amjbot')`) and 0.75 (Fractal geometry model prediction for 3D organisms; `r myCite(c('west1997science', 'west1999science'))`) at 10 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's10_97.5'], 2)`), but that are significantly shallower than 0.75 at 25 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's25_97.5'], 2)`). Microcionidae presents exponents that only overlap 0.67 at 10 ˚C, but not at 25 ˚C (95\% CI at 10 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's10_97.5'], 2)`; 95\% CI at 25 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's25_97.5'], 2)`). On the other hand, for 3D organisms, *Bugula neritina* presents scaling exponents that overlap both 0.67 and 0.75 at 25 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's25_97.5'], 2)`), but are significantly shallower at 10 ˚C (95\% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's10_97.5'], 2)`). Finally, *Bugula stolonifera* presents values that overlap both 0.67 and 0.75 at both temperatures (95\% CI at 10 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's10_97.5'], 2)`; 95\% CI at 25 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's25_97.5'], 2)`).

@@

Here, we show experimentally that the mass-scaling of metabolic rates changes between two temperatures, and in different ways across different species of marine invertebrates. We build on recent empirical evidence documenting similar changes (e.g. `r myCite('killen2010ele')`), yielding further support to the need for new metabolic theories (or reformulation of current ones) that formally incorporate interactive effects between mass scaling and temperature dependence of biological rates. Our results, however, do not indicate consistent mass-temperature interactions across growth forms -- we found idiosyncratic effects of temperature on scaling exponents across closely related species with the same growth form. Possibly, another important axis involved, which has not been considered here, is the acclimation capacity of species upon acute temperature changes. Our experimental set up, however, acclimated organisms for periods (48 h) that seem to be long enough to buffer these effects `r myCite('rezende2014functecol')`.

That metabolic rate scales at an exponent of less than 1 across all four species (with the exception of *Hippopodina* sp. measured at 10 ˚C) has interesting implications for our understanding of the energetics of growth for different colonial animals. For instance, evidence suggest that water clearance and feeding rates scale either isometrically or allometrically with colony size in sponges `r myCite(c('riisgard1993meps', 'kowalke2000jembe', 'mcmurray2014aquatbiol'))`, and these rates may follow the same mass scaling of metabolic rates (see e.g. `r myCite('thomassen1995meps')`). In contrast, most studies indicate that feeding rates scale isometrically or superlinearly in both encrusting and arborescent bryozoans (e.g. `r myCite(c('okamura1984jembe', 'okamura1985jembe', 'pratt2005meps'))`). Our results and those from other studies `r myCite('hartikainen2014jeb')` indicate that metabolic rates scale allometrically in arborescent 3D bryozoans. Consequently, while every unit increase in size results in proportional increase in feeding rate, every unit increase in size results in a less-than-proportional increase in metabolic rate. Thus, as colonies increase in size, their capacity to capture food increases more quickly than the rate at which energy is expended.

This differential between ingestion rates and metabolic costs may explain the observed exponential growth and reproduction observed in some colonial marine invertebrates (e.g. `r myCite(c('marshall2013evolution', 'padilla1996limn', 'winston2010icb'))`) -- because the ratio of intake to expenditure increases with size, larger colonies can allocate proportionately more energy to growth and/or reproduction. In other words, for these colonial marine invertebrates, larger colonies are more efficient at allocating food intake to somatic growth and/or reproduction than smaller colonies. Isometric energy gains and allometric energy expenditure in colonial marine invertebrates may explain why they are often competitively dominant, outgrowing and excluding unitary organisms (which generally present allometric scaling of both feeding and metabolic rates, see e.g. `r myCite(c('kooijman2009book', 'kearney2012amnat'))`) in a range of systems `r myCite(c('buss1980pnas', 'buss1990tree'))`.
